Robust Control of a Fluxonium Qubit

### 2020-11-30 paper/1introduction.tex
updated figures, lots of proofreading

\section{Introduction}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{vandersypen2005nmr, kehlet2004improving, khaneja2005optimal,
  maximov2008optimal, nielsen2010optimal, skinner2003application, tosner2009optimal},
and applications now include superconducting circuits \cite{abdelhafez2020universal,
  chakram2020multimode, fisher2010optimal, gokhale2019partial,
  huang2020engineering, leng2019robust, leung2017speedup, li2020fast,
  xu2020nonadiabatic},
neutral atoms and ions \cite{brouzos2015quantum,
  de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
  treutlein2006microwave, van2016optimal},
nitrogen-vacancy centers in diamond \cite{chou2015optimal,
  dolde2014high, geng2016experimental,
  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
For quantum computation,
optimal control techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system to deviate
from the model used in optimization, leading
to poor experimental performance.
Robust control techniques improve upon
standard optimal control techniques by encoding
model uncertainties
in optimization objectives, yielding performance
guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
We adapt robust control techniques from the robotics community to mitigate
realistic model uncertainties for
a superconducting fluxonium qubit.

QOC techniques have had tremendous success in
engineering high fidelity quantum gates \cite{
  chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
  grace2007optimal, heeres2017implementing, huang2014optimal,
  kelly2014optimal, leng2019robust, liebermann2016optimal,
  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
  spiteri2018quantum, sporl2007optimal}. 
While standard QOC techniques can predict system behavior
with high accuracy, they are sensitive to experimental
errors such as parameter drift, noise, finite control 
resolution, and decoherence. Multiple techniques have been developed to address
these shortcomings. Analytic techniques to mitigate parameter deviation errors
include composite pulses \cite{cummins2000use, cummins2003tackling,
  kupce1995stretched, merrill2014progress},
dynamic and geometric phase considerations
\cite{han2020experimental, xu2020nonadiabatic},
and the DRAG scheme \todo{DRAG citations}.
To mitigate decoherence, Floquet techniques have been employed
\cite{huang2020engineering, mundada2020floquet}.
Numerical techniques to mitigate parameter deviation errors
include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
  wittler2020integrated} and open-loop methods \cite{
  allen2019robust, carvalho2020error, reinhold2019controlling,
  rembold2020introduction, kosut2013robust}.
Numerical techniques to mitigate decoherence include
modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.

In this work, we study three robust control techniques that
make the system's state trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
  \item A sampling method, similar to the work of \cite{allen2019robust,
    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
    to compute derivative information of parameter deviations with respect to the 
    quantum state trajectory.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits,
in particular, taking advantage of the $T_{1}$-dependence of the controls
to mitigate longitudinal relaxation type decoherence and
performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.

>>>>

\section{Introduction}
\todo{Ask Dave to review paragraph}
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{vandersypen2005nmr, kehlet2004improving, khaneja2005optimal,
  maximov2008optimal, nielsen2010optimal, skinner2003application, tosner2009optimal},
and applications now include superconducting circuits \cite{abdelhafez2020universal,
  chakram2020multimode, fisher2010optimal, gokhale2019partial,
  huang2020engineering, leng2019robust, leung2017speedup, li2020fast,
  xu2020nonadiabatic},
neutral atoms and ions \cite{brouzos2015quantum,
  de2008optimal, goerz2011quantum, guo2019high, jensen2019time,
  larrouy2020fast, omran2019generation, rosi2013fast, sorensen2019qengine,
  treutlein2006microwave, van2016optimal},
nitrogen-vacancy centers in diamond \cite{chou2015optimal,
  dolde2014high, geng2016experimental,
  nobauer2015smooth, poggiali2018optimal, rembold2020introduction, tian2019optimal},
and Bose-Einstien condensates \cite{amri2019optimal, sorensen2018quantum}.
For quantum computation,
optimal control techniques are employed to achieve high fidelity gates
while adhering to experimental constraints.
Experimental errors may cause the system to deviate
from the model used in optimization, leading
to poor experimental performance.
Robust control techniques improve upon
standard optimal control techniques by encoding
model uncertainties
in optimization objectives, yielding performance
guarantees over a range of parameters \cite{Zhou97,Morimoto00,Manchester18}.
We adapt robust control techniques from the robotics community to mitigate
parameter deviation errors for
a superconducting fluxonium qubit.

QOC techniques have had tremendous success in
engineering high fidelity quantum gates \cite{
  chou2015optimal, dolde2014high, egger2013optimized, egger2014adaptive,
  grace2007optimal, heeres2017implementing, huang2014optimal,
  kelly2014optimal, leng2019robust, liebermann2016optimal,
  nebendahl2009optimal, rebentrost2009optimal, rebentrost2009optimal2,
  spiteri2018quantum, sporl2007optimal}. 
While standard QOC techniques can predict system behavior
with high accuracy, they are sensitive to experimental
errors such as parameter drift, noise, finite control 
resolution, and decoherence. Multiple techniques have been developed to address
these shortcomings. Analytic techniques to mitigate parameter deviation errors
include composite pulses \cite{cummins2000use, cummins2003tackling,
  kupce1995stretched, merrill2014progress},
dynamic and geometric phase considerations
\cite{han2020experimental, xu2020nonadiabatic},
and the DRAG scheme \cite{motzoi2009simple}.
To mitigate decoherence, Floquet techniques have been employed
\cite{huang2020engineering, mundada2020floquet}.
Numerical techniques to mitigate parameter deviation errors
include closed-loop methods \cite{egger2014adaptive, feng2018gradient, huang2020engineering,
  wittler2020integrated} and open-loop methods \cite{
  allen2019robust, carvalho2020error, reinhold2019controlling,
  rembold2020introduction, kosut2013robust}.
Numerical techniques to mitigate decoherence include
modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.

In this work, we study three robust control techniques that
make the system's state trajectory less sensitive
to static and time-dependent parameter deviations:
\begin{enumerate}
  \item A sampling method, similar to the work of \cite{allen2019robust,
    carvalho2020error, reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in the 
    state estimation community \cite{howell2020direct, julier2004unscented,
      lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
    to compute the sensitivity of the quantum state trajectory
    to deviant parameters.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems for fluxonium-based qubits,
in particular, taking advantage of the $T_{1}$-dependence of the controls
to mitigate longitudinal relaxation type decoherence and
performing phase gates in arbitrary times.
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
utilize an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}.

### 2020-11-30 paper/2background.tex
updated figures, lots of proofreading

\section{QOC + ALTRO \label{sec:background}}
In this section we
review the quantum optimal control problem statement
and introduce the ALTRO solver.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}
  \label{eq:tdse}
\end{equation}
The Hamiltonian has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. To make the problem numerically tractable,
the controls and quantum state are discretized into $N$ knot points (time steps).
The optimization problem is to find the
controls that minimize a functional $J(u)$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
In general, $J$ is a linear combinaion of cost functions on the state
as well as cost functions on the controls \cite{leung2017speedup}.
Typical quantum optimal control
algorithms employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{machnes2015gradient, khaneja2005optimal, leung2017speedup, goerz2019krotov}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
$u_k \in \R^m$ are the augmented state and input augmented control variables,
$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
$h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

@@

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
of penalty weights, $\mu$, whose magnitudes depend on whether the corresponding constraint is active or inactive.
For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR,
the Lagrange multiplier estimates are updated according to,
\begin{equation}
	\lambda \gets \lambda + \mu c(z) ,
\end{equation}
the penalty terms are updated, and the process repeats until convergence.

@@

As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
for imposing arbitrary, simultaneous constraints
on both the state trajectory and the controls and does not restrict
optimization to the constraint manifold. First-order solvers like GRAPE
rely on unconstrained optimization to satisfy desiderata on the state trajectory,
such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
and smoothing the controls below a threshold. The unconstrained approach relies
on a careful choice of competing penalty weights, which is cumbersome and often intractable.
To impose constraints on the controls, such as amplitude restrictions, first-order solvers
employ projected gradient descent methods which restrict optimization to the constraint manifold and
may hinder convergence to the optimal solution \cite{clarkson2010coresets,
  hauswirth2016projected}.
The ALTRO solver delays projecting onto the constraint manifold until the ALM
solve is already at a coarse tolerance, achieving machine-precision satisfaction
for constraints on the controls and the state trajectory. These advantage are critical
for this work, where multiple constraints must be satisfied in addition
to minimizing auxiliary metrics.

>>>>

\section{QOC + ALTRO \label{sec:background}}
In this section we
review the QOC problem statement
and introduce the ALTRO solver.
QOC concerns the evolution of
a (quantum) state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}
  \label{eq:tdse}
\end{equation}
The Hamiltonian has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so-called because they are the means the experimentalist has to
act on the system. To make the problem numerically tractable,
the controls and quantum state are discretized into $N$ knot points (time steps).
The optimization problem is to find the
controls that minimize a functional $J(u)$.
In the simplest case, the functional is
the infidelity between the initial state evolved
to the final knot point and the desired final state
$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
In general, $J$ is a linear combination of cost functions on the state
as well as cost functions on the controls \cite{leung2017speedup}.
Standard QOC
solvers employ automatic differentiation
to compute gradients of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{machnes2015gradient, khaneja2005optimal, leung2017speedup, goerz2019krotov}.

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using a variety of specialized solvers developed by the robotics community
\cite{Schulman13,Tedrake16,Hereid2017FROST,howell2019altro}.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^l$ and
$u_k \in \R^m$ are the augmented state and input augmented control variables,
$f(x_k,u_k)$ is the discrete dynamics function, and $g_k(x_k,u_k)$ and
$h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

@@

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a locally stabilizing closed-loop control policy about the optimal trajectory. However,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^{l + m} \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ is a Lagrange multiplier estimate, and $I_\mu$ is a diagonal matrix
of penalty weights, $\mu$, whose magnitudes depend on whether the corresponding constraint is active or inactive.
For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR,
the Lagrange multiplier estimates are updated according to,
\begin{equation}
	\lambda \gets \lambda + \mu c(z) ,
\end{equation}
the penalty terms are updated, and the process repeats until convergence.

@@

\todo{Ask Dave to review paragraph}
As opposed to standard QOC solvers, the ALTRO solver provides a unified framework
for imposing arbitrary, simultaneous constraints
on both the state and the controls and does not restrict
optimization to the constraint manifold. First-order solvers like GRAPE
rely on unconstrained optimization to satisfy desiderata on the state,
such as achieving a desired gate fidelity, forbidding occupations of particular quantum states,
and smoothing the controls below a threshold. The unconstrained approach relies
on a careful choice of competing penalty weights, which is cumbersome and often intractable.
To impose constraints on the controls, such as amplitude restrictions, first-order solvers
employ projected gradient descent methods which restrict optimization to the constraint manifold and
may hinder convergence to the optimal solution \cite{clarkson2010coresets,
  hauswirth2016projected}.
The ALTRO solver delays projecting onto the constraint manifold until the ALM
solve is already at a coarse tolerance, achieving machine-precision satisfaction
for constraints on the controls and the state. These advantages are critical
for this work, where multiple constraints must be satisfied in addition
to minimizing auxiliary metrics.

### 2020-11-30 paper/3fluxonium.tex
updated figures, lots of proofreading

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the accurate
two-level approximation of its Hamiltonian makes
quantum optimal control on a classical computer inexpensive.
In the two-level
approximation the Hamiltonian takes the form:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a(t)$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
Hamiltonian parameter deviations. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
  {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{(x_1, x_N \; \textrm{given})}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
\end{mini!}
where the augmented state and controls are given by:
\begin{equation}
  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The initial and final conditions are determined by the
desired gate, which requires two quantum states
to specify $\psi^{(1)}$, $\psi^{(2)}$.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
given in \cite{leung2017speedup}:
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
  \label{eq:isomorphism}
\end{equation}
The discrete dynamics function \eqref{eq:dyn_con} integrates the TDSE dynamics for the quantum states
using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
Including objectives at each knot point smoothens the optimization landscape, though
the importance of the final state objective is encoded in the relative
weight $Q_{N} \sim N \cdot Q_{k}$.
We choose $Q$ to be diagonal because it is computationally efficient. This corresponds
to penalizing phase differences between the quantum states, although
other phase-insensitive metrics such as infidelity may be employed.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\psi_{N} = \psi_{f}$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition, the norm of the quantum states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
of subsets from $x_{k}$ is implied in
the constraint equations.

>>>>

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following, we study
the QOC problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the accurate
two-level approximation of its Hamiltonian makes
QOC on a classical computer inexpensive.
In the two-level
approximation the Hamiltonian takes the form:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a(t)$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
Hamiltonian parameter deviations. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k}
    + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{} \label{eq:costfun}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{(x_1 \; \textrm{given})}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
\end{mini!}
The augmented state and augmented controls are:
\begin{equation}
  x = \begin{bmatrix} \psi^{1} \\ \psi^{2} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
  \label{eq:astatecontrols}
\end{equation}
The initial states $\psi^{1}_{1}$, $\psi^{2}_{1}$
are basis vectors of the Hilbert space $\ket{0}, \ket{1}$.
The final states $\psi^{1}_{f}$, $\psi^{2}_{f}$ are the image
of the initial states under the desired gate.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
given in \cite{leung2017speedup},
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
  \label{eq:isomorphism}
\end{equation}
The discrete dynamics function \eqref{eq:dyn_con} integrates the states
using \eqref{eq:tdse} and \eqref{eq:hamiltonian}
and integrates the moments of the flux amplitude.
Exposing lower order moments of the flux amplitude allows us to penalize their norms
in \eqref{eq:costfun}, smoothing the flux amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric.
Including objectives at each knot point smoothens the optimization landscape, though
the importance of the final augmented state objective is encoded in the relative
weight $Q_{N} \sim N \cdot Q_{k}$.
We choose $Q$ to be diagonal because it is computationally efficient. This
metric penalizes phase differences between the states, although
phase-insensitive metrics such as infidelity may be employed.

We impose constraints that
reflect the physical limitations of the apparatus and 
constraints that improve the experimental realization of the control pulse.
To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a_{1} = a_{N} = 0$.
Furthermore, we require $\int a_{N} \ dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a_{k} \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require
$\psi_{N} = \psi_{f}$.
The zero net flux and final state constraints
are satisfied if the final augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition, the norm of the states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$. The selection
of subsets from $x_{k}$ is implied in
the constraint equations.
